---
title: "test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(dplyr)
library(ggplot2)
```

## Reading the Data

The dataset categorizes favorite restaurants based on name,location, cuisine type,  favorite food, cost, date of last visit, and number of visits per year.
```{r}
mine <- read_csv("C:\\Users\\sb3mw\\Desktop\\fav_restaurants.csv")
```
This plot shows the relationship between the cost of meals and how many times per year the restaurant was visited.
```{r}
 ggplot(mine, aes(yearly_visits, cost_per_person)) +
  geom_point()
```

This plot shows the relationship between the cost of meals and the city and state where the restaurant is located.
```{r}
ggplot(mine, aes(cost_per_person, location)) +
  geom_point()
```

This plot shows the relationship between the cuisine type and cost per person of each meal.
```{r}
ggplot(mine, aes(cuisine, cost_per_person)) +
  geom_point()
```


```


## Plotting the Data

