---
title: "Lab 10: Filtering Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
theme_set(theme_minimal())
```

## Pick a dataset

Today you will again have a choice of three datasets to work with. You can
choose either to investigate the daily team ELO ratings in the NBA:

```{r}
nba <- read_csv("https://statsmaths.github.io/stat_data/nba_elo_daily.csv")
```

Metadata from US senator's tweets:

```{r}
tweets <- read_csv("https://statsmaths.github.io/stat_data/senator_tweets_meta.csv")
```

Or swear words given in the movies of Quentin Tarantino:

```{r}
tarantino <- read_csv("https://statsmaths.github.io/stat_data/tarantino.csv")
```

Pick a dataset (feel free to delete the others from the Rmd file if you would
like), and work on the instructions in the next section.

## Instructions

This lab is similar to the last. You'll pick a dataset and then generate an
analysis of that data using data visualizations. The difference here is that
instead of a single plot, I would like you to construct 3-4 plots that
together tell a linear story. Each plot should be separated by a sentence or
two describing what the viewer should take away from the plot. Try to keep the
plots similar in some way; perhaps each simply highlights a different subset
of the data but has the same underlying layers.

Notice that each of the datasets for today are larger than you will probably
be able to use. You will need to filter the data to a particular team, year,
subset of curse words. You may further filter the data to highlight an even
smaller subset of the data within each plot.

## Analysis

```{r}
tarantino_fuck <- filter(tarantino, word == "fuck")
```

```{r}
ggplot(tarantino_fuck, aes(movie)) +
  geom_bar(aes(fill=word))
```
This plot shows the frequency of the word "fuck" in each movie and in which movies it is the most frequent. This particular word appears most frequently in "Pulp Fiction."

```{r}
tarantino_shit <- filter(tarantino, word == "shit")
```

```{r}
ggplot(tarantino_shit, aes(movie)) +
  geom_bar(aes(fill=word))

```
This plot shows the frequency of the word "shit" in each movie and in which movie it is the most frequent. The word "shit" appears most frequently in the movie "Pulp Fiction."


```{r}
tarantino_dick <- filter(tarantino, word == "dick")
```

```{r}
ggplot(tarantino_dick, aes(movie)) +
  geom_bar(aes(fill=word))
```
This plot shows the frequency of the word "dick" in each movie in the dataset. This word only appears in the movies "Jackie Brown" and "Reservoir Dogs" and is said significantly more in "Reservoir Dogs."




