---
title: "Lab 19: Inference for Several Means"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Burrito synergy in SoCal

Today we will look at a dataset of reviews of burritos in southern
California. This is a data set collected by a group of college friends who
live in the greater San Diego area.

```{r}
burrito <- read_csv("https://statsmaths.github.io/stat_data/burrito.csv")
```

The available variables in the data are:

- location: the name of the restaurant
- cost: total cost for the burrito
- yelp: the average Yelp review score
- google: the average Google review score
- chips_included: equals 1 if chips are included with the burrito
- hunger: score from 1 to 5 on how much the burrito filled up the reviewer
- tortilla: rating from reviewer; 1 (terrible) to 5 (awesome)
- temp: rating from reviewer; 1 (terrible) to 5 (awesome)
- meat: rating from reviewer; 1 (terrible) to 5 (awesome)
- fillings: rating from reviewer; 1 (terrible) to 5 (awesome)
- meat_filling: rating from reviewer; 1 (terrible) to 5 (awesome)
- uniformity: rating from reviewer; 1 (terrible) to 5 (awesome)
- salsa: rating from reviewer; 1 (terrible) to 5 (awesome)
- synergy: rating from reviewer; 1 (terrible) to 5 (awesome)
- wrap: rating from reviewer; 1 (terrible) to 5 (awesome)
- overall: rating from reviewer; 1 (terrible) to 5 (awesome)
- lat: latitude (in degrees North) of the taco restaurant
- lon: longitude (in degrees East) of the taco restaurant

Fit a linear regression predicting the average yelp score of locations in this
dataset and look at the model using `reg_table`:

```{r}
model<-lm_basic(yelp~1, data=burrito)
reg_table(model, level=0.95)

```


The average Yelp score of all restaurants in this area is just 3.7; looking
at the confidence interval from the model does it seem likely that burrito
locations have a higher Yelp rating than locations in general, or is this
just noise in the data?

**Answer**: It is likely that burrito locations have a higher Yelp rating than locations in general because the range doesn't go below 3.9.

The regression model confidence interval requires that the data are sampling
independently from the larger population. Why might this dataset not hold to
this assumption?

**Answer**: Because the restaurants were picked beforehand they most likely picked the places that were already known to be the best ones, resulting in high scores.

Now fit a linear regression that predicts the yelp score as a function of
whether the burrito includes chips.

```{r}
model<-lm_basic(yelp~1 + chips_included, data=burrito)
reg_table(model, level=0.95)
```

Describe what the slope coefficient in the previous model really means
in terms of the data:

**Answer**: The overall score of the burrito is approximately .05% higher if chips are included in the order.

Is the slope coefficient statistically significant (compared to zero)?

**Answer**: Yes

Describe this in terms of your answers above.

**Answer**: This is because if it was zero whether or not chips are included would have no impact on the overall score, but because it is above zero it does have an impact.


The word *synergy* can be defined as:

> the interaction or cooperation of two or more organizations, substances,
> or other agents to produce a combined effect greater than the sum of
> their separate effects.

This is not something I would have thought of when considering tacos, but
let's see how synergy effected the scores of both the reviewers as well as
Google and Yelp users in general.

First, I'll add a variable called high_synergy defined as whether the synergy
score is greater than 3.5.

```{r}
burrito$high_synergy <- (burrito$synergy > 3.5)
```

Fit a regression predicting the Yelp score as a function of the
variable `high_synergy`.

```{r}
model<-lm_basic(yelp~1 + high_synergy, data=burrito)
reg_table(model, level=0.95)
```

Is there strong evidence that Yelp scores are higher for high synergy burrito
restaurants?

**Answer**: No, because the range of confidence goes from negative to positive so it is impossible to conclude that Yelp scores are higher for high synergy burrito restaurants.

Now fit a regression predicting the overall score as a function of the
variable high_synergy.

```{r}
model<-lm_basic(overall~1 + high_synergy, data=burrito)
reg_table(model, level=0.95)
```

Is there evidence that high synergy burrito locations have a higher overall
score than low synergy burrito locations?

**Answer**: Yes, because the confidence intervals are both positive so there is enough evidence that high synergy burrito locations have a higher overall score.

Using the model from the previous question, what does the model predict
will be the overall score of a location with low synergy?

**Answer**: 2.6963

What does the model predict will be the overall score of a location with
high synergy?

**Answer**: 3.8641

Finally, fit a regression model that predicts the overall score as a function
of the variable `synergy` coded as a factor.

```{r}
model<-lm_basic(overall~1 + factor(synergy), data=burrito)
reg_table(model, level=0.95)
```

What does the model predict the overall score for a restaurant with a
synergy score of 3 will be?

**Answer**: 0.6917





