---
title: "Relationships Between Race, Commute, and Income in Denver, CO"
author: "Sofia Barnes-Horowitz"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(ggmap)
```


```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")

denver <- filter(tracts, cbsa=="Denver-Aurora-Lakewood, CO")
```

##Background:
The city of Denver, often referred to as the "Mile High City," and the suburbs surrounding it are known for their beautiful mountain views and a short but rich cultural history.  However, it is also known for not being particularly diverse, whether looking at factors of ethnicity, income or education level, number of languages spoken, religion, or household type.  In fact, Denver ranked 36th in the nation for diversity (Sevits, 2017).  And in terms of commuting, approximately 85% of commuters prefer to drive, while only 4% use public transportation, and the average commute time is 27.2 minutes, only slightly more than the national average (Hearsum 2017).

##Introduction
With these figures in mind, when looking at the relationship between census data and a large metropolitan area such as Denver, Colorado, it is interesting to compare factors such as commute methods, times, and income level with factors such as race; specifically the dichotomy between white and black residents.  Before constructing the data, the assumed relationship would be that wealthy, white residents live in the suburbs, resulting in a longer car commute, and lower-income minority residents (in this case measuring percentages of African Americans) would live closer to the city, resulting in a shorter commute and one not necessarily by car, but by public transportation, walking, biking, etc.  It is important to note, however, that the public transportation system in Denver, while very effective, is extremely new in comparison to those of bigger cities such as New York, Boston, or Chicago, which could potentially skew the results for percentage of residents taking cars in the graphics below (Hearsum, 2017).  Also, as stated above, the entire metropolitan area is not as racially diverse as many other cities, so it is logical that there would be much larger majorities of white people when looking at each variable and the relationships between them.


```{r}
ggplot(denver, aes(commute_car, race_white)) +
  geom_point(aes(color=race_black)) +
  scale_color_viridis() +
  ggtitle("Relationship Between Race and Commute By Car") +
  xlab("Commute by Car") +
  ylab("White Commuters")
```

The first relationship is between white commuters and the percentage of those that commute by car, with a color scale denoting numbers of black commuters that are present in the dataset.  As predicted, the vast majority of residents prefer to commute by car, and the vast majority of commuters are white, as evidenced by the clump of purple in the top right corner.  However, it is interesting to note that the method of transportation does not seem to be affected by race, as the black and mixed-raced residents follow the same trend as the white residents.  Now that we know that the majority of commuters get to work by car, it is interesting to look at how long the average commute takes.


```{r}
denver$ctime_25_plus <- denver$ctime_25_29 + denver$ctime_30_34 + denver$ctime_35_39 + denver$ctime_40_44 
```

```{r}
ggplot(denver, aes(ctime_25_plus, race_white)) +
  geom_point(aes(color=race_black)) +
  scale_color_viridis() +
  ggtitle("Relationship Between Race and Commute Time Over 25 Minutes") +
  xlab("Commute Over 25 Minutes") +
  ylab("White Commuters")
  
```

For the purposes of this report, it makes sense to define a "short" commute as 20 minutes or less, in which case an average commute, i.e. one from the suburbs into the city, would be approximately between 25 and 45 minutes.  When looking at the graph below, commute times are clustered in the middle, indicating that only about 50 percent of commuters take 25 minutes or more to get to work each day.  Also, just as in the first graph, there is little to no relationship between race and commute time, as the black and mixed-race populations followed the same patterns as the white population.  The final graph, however, shows the relationship between race (again, between white and black residents) and income, in this case those making $75,000 per year or more.  


```{r}
denver$income_075_100_plus <- denver$income_075_100 + denver$income_100_125 + denver$income_125_150 
```

```{r}
ggplot(denver, aes(income_075_100_plus, race_white)) +
  geom_point(aes(color=race_black)) +
  scale_color_viridis() +
  ggtitle("Relationship Between Race and Income") +
  xlab("Residents Making $75,000 Per Year or More") +
  ylab("White")
  

```

The use of looking at incomes of $75,000 or more per year is significant because in the state of Colorado, it takes a minimum of $131,000 per year to be considered upper income, while the median household income is about $44,000 per year (Loudenback and Gould, 2017).  Therefore, an average annual income of $75,000 per year or more indicates an upper middle class status on the lower end of the scale, and significant wealth on the higher end.  In the graph, a little more than 50 percent of the given population is making $75,000 per year or more- while the trend towards the right of the scale is small, it is significant in showing the amount of wealth in the Denver metropolitan area.  In terms of the relationship between income and race, black and mixed-race residents again followed the same trends as their white counterparts.

##Conclusion:
Despite the previous assumption that the data would reveal that white, wealthy residents prefer to commute by car and have a longer commute time, the data instead showed that only one of those assumptions holds true: a vast majority of Denver residents do in fact commute by car.  However, commute times were relatively evenly divided, even after splitting them up into categories of short, average, and long (the graph depicted the "average" of 25 minutes of more), between varieties of times, with no clear trend.  In terms of income, Denver clearly has less of a dichotomy between the lower-income, middle class, and wealthy residents, with a small trend towards the higher-income side of the scale.  This is proven by external data, as the poverty rate in metropolitan Denver has been rising, but at a much slower rate than the rest of the country, which indicates that there is a generally more even distribution of wealth within the city and surrounding suburbs than in other major U.S. cities (Aguilar, 2017).  As briefly mentioned in the introduction, there are a variety of external factors that could have influenced these results: for example, there could be more of a dichotomous relationship between whites and Hispanics as opposed to whites and blacks within each subsequent variable, and again with the variable of race, Denver is an incredibly white city and not very ethnically diverse, so race ultimately reveals little in relation to commute times, income level, and mode of transportation.  Also, the relatively equal dispersion of wealth and commute times could be attributed to a lack of specificity: for example, the data could have been further broken down into commute time and wealth by neighborhood, which would indicate if there was a more direct relationship between the two variables and if race played a role in their relationship, because the entire metropolitan area is extremely large and spans across wealthy and lower-income neighborhoods.  However, it is still interesting to observe the distribution of Denver's white population in transportation preference, average commute time, and average income, as the results would like very different in another American city.

## References
Aguilar, John. "Suburban poverty rising more slowly in Denver..." Denver Post, 14 July 2017, https://www.denverpost.com/2017/07/14/poverty-suburban-denver-apartment-list-report/.

Hearsum, Katie. "What's it like to live in Denver, CO?" U.S. News & World Report, 2017, https://realestate.usnews.com/places/colorado/denver.

Loudenback, Tanza, and Skye Gould. “How Much You Have to Earn to Be Considered Rich in Every State.” Business Insider, 16 Sept. 2017, www.businessinsider.com/upper-income-us-state-2017-9.

Sevits, Kurt. "How diverse are Colorado's cities compared to others? Not very." Denver7, 4 May 2017, https://www.thedenverchannel.com/news/local-news/how-diverse-are-colorados-cities-compared-to-others-not-very.



